The SMT Competition 2015--2018

### 2018-10-11 2018/report/SMT-COMP.tex
Work in progress

\section{Introduction}
\label{sec:intro}

Satisfiability Modulo Theories (SMT) is a generalization of Boolean
satisfiability (SAT), the satisfiability decision problem for
propositional logic.  In place of Boolean variables, SMT formulas may
contain terms that are built from function and predicate symbols drawn
from a number of background theories.  Background theories, which are
typically motivated by application domains, include the theory of
arrays~\cite{TODO}, integer and real arithmetic~\cite{TODO},
bit-vectors~\cite{TODO}, and floating-point numbers~\cite{TODO} among
others.  For instance, the following is an SMT formula over the
combination of integer arithmetic and uninterpreted functions:
%
$$x \leq y \wedge y \leq x \wedge P (f(x) - f(y)) \wedge \neg P(0)$$
%
Software tools to determine the satisfiability of such formulas are
called SMT solvers.  With its rich input language, SMT has
applications in software engineering, optimization, and many other
areas~\cite{DeMoura:2011:SMT}.

@@

The International Satisfiability Modulo Theories Competition
(SMT-COMP) is an annual competition between SMT solvers.  It is
affiliated with the International Workshop on Satisfiability Modulo
Theories.  Solvers are submitted to the competition by their
developers, and pitted against each other in a number of tracks and
divisions.  The SMT Competition was part of the Federated Logic
Conference (FLoC) Olympic Games, which comprised 14 competitions in
various areas of computational logic, in Vienna in~2014, and again in
Oxford~(UK) in~2018.  It was last described in a 2014 competition
report~\cite{DBLP:journals/jsat/CokDW14}.  In this paper, we report on
the design and selected results of the competition during the last
FLoC Olympiad, from 2015 to~2018.

@@

The SMT Competition was instituted in 2005.  It has been held annually
since then (with the exception of 2013, when an
evaluation~\cite{DBLP:journals/jar/CokSW15} was conducted instead of a
competition).  The main goals of the competition were to spur adoption
of the common, community-designed SMT-LIB formats (see
Section~\ref{sec:smtlib}), and to spark further advances in SMT,
especially for verification~\cite{smtcomp2005}.  These, together with
providing a useful yardstick for users and developers of SMT solvers,
are still its main goals today.  The competition has been successful
in establishing SMT-LIB as the \emph{de facto} standard language for
SMT solvers.  In recent years, the focus in this regard has shifted
towards promoting some of the newer extensions of the SMT-LIB
language, such as a floating-point theory~\cite{TODO} or algebraic
datatypes~\cite{TODO}.

The competition is organized under the direction of the SMT Steering
Committee, which appoints the competition chair.  This appointment
should happen ``within 30 days from the end of the current edition of
the SMT workshop''~\cite{smt-bylaws} but in practice has often
happened later.  The competition chair then assembles a team of
organizers, and distributes the work for the next competition.
%
The SMT Workshop includes a block of time to present the competitors
and results of the competition.  The workshop is affiliated with a
different conference each year: with CAV in 2015, with IJCAR in 2016,
again with CAV in 2017, and with FLoC in~2018.  Consequently, its date
varies by several weeks.  Competition results were presented on
July~19 in 2015, on July~2 in 2016, on July~23 in 2017, and on July~13
in~2018.

Important deadlines are effectively determined by calculating
backwards from the date of the next SMT Workshop.  The competition's
computational workload should be completed approximately two weeks
before the workshop date, to give participants and other parties a
chance to scrutinize the data for errors, allow partial re-runs of
competition jobs if necessary, and give the organizers time to prepare
the results presentation.  While making all data publicly available
well in advance of the official results presentation may impact the
suspense, it has on several occasions helped to uncover serious
problems with the competition tools or the StarExec framework (see
Section~\ref{sec:starexec}) that could have led to invalid results
otherwise.
\begin{tjark}
  Discuss these issues somewhere, add (forward?) reference here.
\end{tjark}

It takes up to three weeks to run all competition jobs on the StarExec
cluster, so the final submission deadline for solvers needs to be
about five weeks before the workshop date.  The competition imposes a
separate deadline for first versions of solvers about seven weeks
before the workshop date.  This deadline has proved useful for the
organizers to obtain an accurate estimate on the number of competing
solvers, and to run preliminary tests with some of the submitted
solvers to identify potential issues.  Participants can still make
changes to their solver until the final deadline but are encouraged to
use this period for bug-fixing only.

@@

The competition rules are revised each year, and a (near final) draft
version of the rules is made public on the competition web site
typically around mid-April.  The competition chair has ultimate
responsibility for the rules, but changes are often preceded by
discussions among the organizers or within the SMT community.  Such
discussions are initiated by a Call for Comments that is typically
sent out to the SMT-COMP mailing list~\cite{TODO} in February or early
March.  Publicly, this call is the first harbinger of a new edition of
the SMT Competition.

@@

\section{SMT-LIB Logic, Language and Solvers}
\label{sec:context}

@@

Participants must upload their solvers to StarExec (see
Section~\ref{sec:starexec}) to enter them into the competition, and
additionally submit information about each competing solver---such as
the tracks and divisions into which the solver is being entered---to
the competition organizers.  Until 2016, this information was
collected by email.  Since 2017, a web form is being used for this
purpose.  This has greatly reduced the number of incomplete or
ambiguous submissions.

@@

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{13cm}|} %TODO: fix width
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC [\TODO]       & Valera Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE            & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector         & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI           & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo         & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3              & Kshitij Bansal
                        (New York University)\\
    CVC4              & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP          & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink        & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2          & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB              & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B               & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT             & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog            & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT           & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol       & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT        & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP               & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt            & Masahiro Sakai
                        (\TODO)\\
    Vampire           & Giles Reger
                        (University of Manchester)\\
    veriT             & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat              & Martin Velez
                        (University of California, Davis)\\
    Yices2            & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3                & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}  
\end{table}

@@

The competition-wide ranking for the main track was retained and
refined after~2014.  A contentious issue with this ranking has been
the scoring of incorrect solver responses.  In the per-division
rankings, an incorrect response causes a solver to be ranked behind
all correct solvers~\cite{rules18}.  But the competition's main track
consists of many independent divisions.  To balance the strong
interest in correct solvers against the risk of stifling innovation,
the competition-wide ranking was revised in~2017 to use a fixed
numeric penalty for each division with incorrect responses.  The value
of this penalty is set so that entering a (possibly buggy) solver that
can solve all benchmarks into a division has a positive expected score
if the probability of a soundness bug in the solver is below~$20\,\%$.

The 2018 edition of the competition was again part of the FLoC Olympic
Games.  After last-minute negotiations with the Olympic Games
organizers, it was decided that the SMT Competition would award seven
Kurt G\"odel medals.  The 2018 competition organizers conferred these
to the following solvers:
\begin{itemize}
\item The best three main-track solvers according to the
  competition-wide ranking (first place: CVC4, second place:
  Yices-2.6.0, third place: SMTInterpol);
\item the solver(s) that won the QF\_BV division in the main track
  (sequential ranking: Boolector, parallel ranking: Minkeyrink-MT);
\item the solver that won the most competitive divisions in the
  application track (Yices-2.6.0);
\item the solver that won the most competitive divisions in the
  unsat-core track (Yices-2.6.0).
\end{itemize}

In comparison to the competition-wide ranking for the main track, the
metric used for the application and unsat-core track---simply counting
the number of competitive divisions that a solver has won---was fairly
unsophisticated.  Future editions of the competition should perhaps
refine this to use a ranking scheme that is more similar to the
main-track ranking, by weighing divisions according to their relative
importance, and by also rewarding good performance by solvers other
than division winners.  A difficulty in adapting the main-track
ranking for the other tracks is that each main-track benchmark is
worth one point (before weighing) if solved correctly, while
benchmarks in the application and unsat-core track are worth a varying
number of points.  Moreover, in the unsat-core track, the size of the
smallest unsatisfiable core---and hence the number of points
attainable on any given benchmark---may not be known precisely.

>>>>

\section{Introduction}
\label{sec:introduction}

Satisfiability Modulo Theories (SMT) is a generalization of Boolean
satisfiability (SAT), the satisfiability decision problem for
propositional logic.  In place of Boolean variables, SMT formulas may
contain terms that are built from function and predicate symbols drawn
from a number of background theories.  Background theories, which are
motivated by application domains, include the theory of
arrays~\cite{TODO}, integer and real arithmetic~\cite{TODO},
bit-vectors~\cite{TODO}, and floating-point numbers~\cite{TODO} among
others.  For instance, the following is an SMT formula over the
combination of integer arithmetic and uninterpreted functions:
%
$$x \leq y \wedge y \leq x \wedge P (f(x) - f(y)) \wedge \neg P(0)$$
%
Software tools to determine the satisfiability of such formulas are
called SMT solvers.  With its rich input language, SMT has
applications in software engineering, optimization, and many other
areas~\cite{DeMoura:2011:SMT}.

@@

The International Satisfiability Modulo Theories Competition
(SMT-COMP) is an annual competition between SMT solvers.  It was
instituted in~2005~\cite{TODO}, and is affiliated with the
International Workshop on Satisfiability Modulo Theories.  Solvers are
submitted to the competition by their developers, and pitted against
each other in a number of tracks and divisions.  The SMT Competition
was part of the Federated Logic Conference (FLoC) Olympic Games, which
comprised 14 competitions in various areas of computational logic, in
Vienna in~2014, and again in Oxford~(UK) in~2018.  It was last
described in a 2014 competition
report~\cite{DBLP:journals/jsat/CokDW14}.  In this paper, we report on
the design and selected results of the competition during the last
FLoC Olympiad, from 2015 to~2018.

@@

The original goals of the competition were to spur adoption of the
common, community-designed SMT-LIB formats (see
Section~\ref{sec:smtlib}), and to spark further advances in SMT,
especially for verification~\cite{smtcomp2005}.  These, together with
providing a useful yardstick of performance for users and developers
of SMT solvers, are still its main goals to date.  The competition has
been successful in establishing SMT-LIB as the de~facto standard
language for SMT solvers.  In recent years, the focus in this regard
has shifted towards promoting some of the newer extensions of the
SMT-LIB language~\cite{BarFT-RR-17}, such as its floating-point theory
or algebraic datatypes.

The competition is organized under the direction of the SMT Steering
Committee, which appoints the competition chair.  This appointment
should happen ``within 30 days from the end of the current edition of
the SMT workshop''~\cite{smt-bylaws} but in practice has often
happened later.  The competition chair then assembles a team of
organizers, and distributes the work for the next edition of the
competition.
%
The SMT Workshop includes a block of time to present the competitors
and results of the competition.  The workshop is affiliated with a
different conference each year: with CAV in 2015, with IJCAR in 2016,
again with CAV in 2017, and with FLoC in~2018.  Consequently, its date
varies by several weeks.  Competition results were presented on
July~19 in 2015, on July~2 in 2016, on July~23 in 2017, and on July~13
in~2018.

Important competition deadlines are determined by calculating
backwards from the date of the next SMT Workshop.  The competition's
computational workload should be completed approximately two weeks
before the workshop date.  This gives participants and other parties a
chance to scrutinize the data for errors, allows partial re-runs of
competition jobs if necessary, and gives the organizers time to
prepare the results presentation.  While making all data publicly
available well in advance of the official results presentation may
impact the suspense, it has on several occasions helped to uncover
serious problems with the competition tools or the StarExec framework
(see Section~\ref{sec:starexec}) that could have led to invalid
results otherwise.

It typically takes up to three weeks to run all competition jobs on
the StarExec cluster, so the final submission deadline for solvers
needs to be about five weeks before the workshop date.  The
competition imposes a separate deadline for first versions of solvers
about two weeks before the final submission deadline.  This earlier
deadline has proved useful for the organizers to obtain an accurate
estimate on the number of competing solvers, and to run preliminary
tests with some of the submitted solvers to identify potential issues.
Participants can still make changes to their solver until the final
deadline but are encouraged to use this period for bug-fixing only.

@@

The competition rules are revised each year, and a (near final) draft
version of the rules is made public on the competition web site
usually around mid-April.  The competition chair has ultimate
responsibility for the rules, but changes are often preceded by
discussions among the organizers or within the SMT community.  Such
discussions are initiated by a Call for Comments that is sent out to
the SMT-COMP mailing list~\cite{smtcomp-mailinglist} typically in
February or early March.  Publicly, this call is the first harbinger
of a new edition of the SMT Competition.

@@

\section{The SMT-LIB Language and Library}
\label{sec:smtlib}

All problems used in the SMT Competition come from the SMT Library
(SMT-LIB).  This is a large collection of benchmark problems from
various sources.  The SMT Library is available from a public
repository~\cite{smtlib-repository}, and each new release is mirrored
on StarExec (see Section~\ref{sec:starexec}), the cluster on which the
competition runs.

Benchmarks in the SMT Library are written in the SMT-LIB
language~\cite{BarFT-RR-17}, a text-based language that defines the
syntax and semantics of SMT solver input and output.  This includes a
language for terms and formulas, and a command language for
interacting with SMT solvers.  For instance,
Figure~\ref{fig:smtlib-example} shows a possible representation of the
example formula from Section~\ref{sec:introduction} as a benchmark in
SMT-LIB syntax.

\begin{figure}
\begin{verbatim}
foo
\end{verbatim}
\caption{A benchmark problem in SMT-LIB syntax.}
\label{fig:smtlib-example}
\end{figure}

@@

Participants must upload their solvers to StarExec to enter them into
the competition, and additionally submit information about each
competing solver---such as the tracks and divisions into which the
solver is being entered---to the competition organizers.  Until 2016,
this information was collected by email.  Since 2017, a web form is
being used for this purpose.  This has greatly reduced the number of
incomplete or ambiguous submissions.

@@

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{13cm}|} %TODO: fix width
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC [\TODO]       & Valera Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE            & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector         & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI           & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo         & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3              & Kshitij Bansal
                        (New York University)\\
    CVC4              & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP          & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink        & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2          & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB              & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B               & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT             & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog            & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT           & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol       & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT        & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP               & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt            & Masahiro Sakai
                        (Toshiba Corporation)\\
    Vampire           & Giles Reger
                        (University of Manchester)\\
    veriT             & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat              & Martin Velez
                        (University of California, Davis)\\
    Yices2            & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3                & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}  
\end{table}

@@

The competition-wide ranking for the main track was retained and
refined after~2014.  A contentious issue with this ranking has been
the scoring of incorrect solver responses.  In the per-division
rankings, an incorrect response causes a solver to be ranked behind
all correct solvers~\cite{rules18}.  But the competition's main track
consists of many independent divisions.  To balance the strong
interest in correct solvers against the risk of stifling innovation,
the competition-wide ranking was revised in~2017 to use a fixed
numeric penalty for each division with incorrect responses.  The value
of this penalty is currently set so that entering a (possibly buggy)
solver that can solve all benchmarks into a division has a positive
expected score if the probability of a soundness bug in the solver is
below~$20\,\%$.

The 2018 edition of the competition was again part of the FLoC Olympic
Games, which sponsored seven Kurt G\"odel medals.  The competition
organizers awarded these to the following solvers:
\begin{itemize}
\item The best three main-track solvers according to the
  competition-wide ranking (first place: CVC4, second place:
  Yices-2.6.0, third place: SMTInterpol);
\item the solver(s) that won the QF\_BV division in the main track
  (sequential ranking: Boolector, parallel ranking: Minkeyrink-MT);
\item the solver that won the most competitive divisions in the
  application track (Yices-2.6.0);
\item the solver that won the most competitive divisions in the
  unsat-core track (Yices-2.6.0).
\end{itemize}

In comparison to the competition-wide ranking for the main track, the
metric used for the application and unsat-core track---simply counting
the number of competitive divisions that a solver has won---was rather
unsophisticated.  Future editions of the competition should perhaps
refine this to use a ranking scheme that is more similar to the
main-track ranking, by weighing divisions according to their relative
importance, and also by rewarding good performance from solvers other
than division winners.  A difficulty in adapting the main-track
ranking for the other tracks is that each main-track benchmark is
worth one point (before weighing) if solved correctly, while
benchmarks in the application and unsat-core track are worth widely
varying numbers of points.  Moreover, in the unsat-core track, the
size of the smallest unsatisfiable core---and hence the number of
points attainable on any given benchmark---may not be known precisely.
